SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-03 docs/questioning.tex
runtime tainting draft ready

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (locus {\sc y}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protect the web user from exploitations
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions to
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (locus {\sc x}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt eval} and {\tt write} that are
used to assemble innocuous-looking parts into harmful
strings would fail. A critical issue regarding
the above frameworks is that they cannot
define policies that restrict the behavior of third-party
scripts introduced by script inclusion and {\tt iframe}
tags. For instance, consider a web page ({\tt foo.net})
that prints the value of the a query parameter ({\tt query})
from the page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe}
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls ({\tt malicious.com}).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

There are different implementations of these approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}). However, this is a simplistic
approach that may lead to many false positives.
{\sc csse}~\cite{PB05} and the system by {\it Xu et al.}~\cite{XBS06}
apply further checks when they identify that tainted data
have reached a sink. In particular, {\it Xu et al.}~\cite{XBS06}
tracks taint information at the level of bytes in memory.
To distinguish between legitimate and malicious uses
of untrusted data that reach a sink,
they search the data for suspicious symbols by using
regular expressions. {\sc csse} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and other. When tainted data reaches a sink {\sc csse} performs
syntactic checks it based on the aforementioned metadata
(locus {\sc z}). {\it Xu et al.}~\cite{XBS06}
have proposed a 
In the case of {\sc dsl}-driven injection attacks,
the above implementations can be efficient. However, this
does not apply in the case of JavaScript injection.
Currently, with the wide use of JavaScript these mechanisms
would produce numerous false alarms because instead
of checking for tainted variables ``in" scripts,
they look inside {\sc html} responses ``for" scripts.

% As the author of {\sc csse} state~\cite{PB05}, the prevention
% of {\sc xss} attack requires a more complex analysis.

@@

The training subcategory, includes some mechanisms
that can be easily circumvented. For example,
consider an application that will send the password 
for a forgetful user via email by executing
the following query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
In a similar manner, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a mimicry attack like the
one mentioned in Subsubsection {\sc tba}.

@@

@@

  &   {\it CsFire}~\cite{DDHPJ10} & & & {\bf ?} & {\sc csrf} \\

@@

	\multirow{8}{*}{Runtime Tainting}

@@

  \multirow{3}{*}{Hybrid}

@@

>>>>

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (locus {\sc y}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protect the web user from exploitations
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions to
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (locus {\sc x}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt eval} and {\tt write} that are
used to assemble innocuous-looking parts into harmful
strings would fail.

A critical issue regarding
the above frameworks is that they cannot
define policies that restrict the behavior of third-party
scripts introduced by script inclusion and {\tt iframe}
tags. For instance, consider a web page ({\tt foo.net})
that prints the value of the a query parameter ({\tt query})
from the page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe}
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls ({\tt malicious.com}).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

There are different implementations of these approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (locus {\sc z}).
However, this is a simplistic approach that may
lead to many false positives.

The following schemes apply further checks when
they identify that tainted data have reached a sink.
In particular, {\it Xu et al.}~\cite{XBS06}
tracks taint information at the level of bytes in memory.
To distinguish between legitimate and malicious uses
of untrusted data that reach a sink,
they search the data for suspicious symbols by using
regular expressions. {\sc csse}~\cite{PB05} associates
tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application
and other. When tainted data reaches a sink {\sc csse} performs
syntactic checks it based on the aforementioned metadata
(locus {\sc z}). {\it {\sc php} Aspis}~\cite{PMP11} works
in a similar way. To obtain metadata, it takes advantage
of specific features provided by {\sc php}.
Note that the above schemes detect potential attacks
in locus {\sc z}. This automatically means that they
cannot detect non-persistent JavaScript-driven
{\sc xss} attacks. {\sc wasc}~\cite{NLC07} 
though, can deal with such attacks because it
analyzes {\sc html} responses to check
if there are tainted data that contain scripts (locus {\sc y}).
A recent study showed that there are ways to circumvent
the majority of the above schemes~\cite{NBR14}.
Furthermore, most of them are hard to deploy since
the majority of input vectors, string operations
and output vectors must be instrumented.

{\it Vogt et al.}~\cite{VFJKKV07}
have developed a tainting scheme that follows a different
approach. Contrary to the aforementioned schemes that
operate on the server side, they track taint sensitive
information on the client side (locus {\sc x}).
This is a form of positive taint tracking, where
taint data is considered to be legitimate.
In essence, they detect JavaScript-driven {\sc xss}
attacks by ensuring that a script can send sensitive
user data only to the site from which it came from.
This though, may lead to false alarms.
{\it Stock et al.}~\cite{SLMS14} propose a scheme
that also operates in the browser (locus {\sc x}).
The scheme focuses on the detection of
{\sc dom}-based {\sc xss} attacks.
In such an attack, the malicious payload is
executed as a result of the modification of the
{\sc dom} environment so that the client-side
code runs in an unanticipated manner even if the
{\sc http} response that triggers the injection seems legitimate.
The scheme is different to the previous one because it
marks and observes data that are considered harmful.
Specifically, it employs a taint-enhanced JavaScript
engine that tracks the flow of attacker-controlled data.
To detect potential attacks the scheme utilizes
{\sc html} and JavaScript parsers that can
identify the generation of code coming from
tainted data.

% @dimitro: move this to the analysis Section:
% This scheme is an efficient scheme that focuses
% on a specific category of {\sc xss} attacks and
% takes into consideration many attack vectors
% that many of the previous schemes do not.
% Also: 
% As the author of {\sc csse} state~\cite{PB05}, the prevention
% of {\sc xss} attack requires a more complex analysis.

@@

The training subcategory, includes some mechanisms
that can be easily circumvented. For example,
consider an application that will send the password 
for a forgetful user via email by executing
the following query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
A traditional {\sc sql} injection attack protection system like {\sc didafit},
could be easily escaped since a malicious user, could obtain the
administrator's password by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.

@@

{\it Diglossia}~\cite{SMS13} performs positive taint tracking and
parse-tree validation.

@@

  &   {\it CsFire}~\cite{DDHPJ10} & 35 & (419582,1141807,0,3)\_r\tnote{3} & {\bf ?} & {\sc csrf} \\

@@

	\multirow{7}{*}{Runtime Tainting}

@@

  \multirow{4}{*}{Hybrid}

@@

    \item[3] The number involve requests.
